---
title: "HATCHER_Skills_Test_3"
output: html_document
date: "2025-04-20"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. Load and clean FacultySalaries_1995.csv file and Re-create the given graph:

```{r include=FALSE}
# load in the needed libraries
library(tidyverse)
library(janitor)
library(ggplot2)
library(broom)

# load in the data
fac_dat <- read.csv("FacultySalaries_1995.csv")
```

First we need to clean the data by taking the three columns AvgFullProfSalary, AvgAssocProfSalary, AvgAssistProfSalary and assigning their names to Rank and values to Salary (make the data set longer).
```{r}
faculty_long <- fac_dat %>% 
  pivot_longer(
    cols = c(AvgFullProfSalary, AvgAssocProfSalary, AvgAssistProfSalary),
    names_to = "Rank",
    values_to = "Salary") %>% 
  mutate(Rank = case_when(
    Rank == "AvgFullProfSalary" ~ "Full",
    Rank == "AvgAssocProfSalary" ~ "Assoc",
    Rank == "AvgAssistProfSalary" ~ "Assist",
    TRUE ~ Rank))
```
  
The variables rank, tier, and state need to be converted into factors for plotting.
```{r}
faculty_long$Rank <- as.factor(faculty_long$Rank)
faculty_long$Tier <- as.factor(faculty_long$Tier)
faculty_long$State <- as.factor(faculty_long$State)
```

Now we create a boxplot faceted by tier with the x-axis as rank, y-axis as salary, and fill as rank.
```{r}
faculty_long_plot <- filter(faculty_long, Tier != "VIIB")

ggplot(faculty_long_plot, aes(x = Rank, y = Salary, fill = Rank)) +
  geom_boxplot() +
  facet_wrap(~ Tier) +
  labs(y = "Salary", x = "Rank") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 60))
```

## 2. Build an ANOVA model and display the summary output.
The model tests the influence of state, tier, and rank on salary but does not include any interactions between the predictors.

```{r}
anova_mod <- aov(data = faculty_long, Salary ~ State + Tier + Rank)

summary(anova_mod)
```

Based on the ANOVA model above, we can conclude that all three variables (State, Tier, and Rank) have effects on faculty salaries that are statistically significant (the chance that these differences occurred due to chance is incredibly unlikely).  
  
Since Rank has the highest sum of squares and the highest F-value, we can conclude that Rank is the most important variable in explaining salary.  

Although all three variables are significant, which state a professor works in matters the least compared to their rank or university tier.  
  
  
## 3. Load and look at "Juniper_Oils.csv" data and then tidy it.

We first need to load in the data set Juniper_Oils.csv and take a look at the structure. 
```{r}
juniper_dat <- read.csv("Juniper_Oils.csv")

str(juniper_dat)
```

Now we need to make a vector of the chemical compound names and then tidy up the chemical data while keeping the rest.
```{r}
chemicals <- c("alpha.pinene","para.cymene","alpha.terpineol","cedr.9.ene",
               "alpha.cedrene","beta.cedrene","cis.thujopsene","alpha.himachalene",
               "beta.chamigrene","cuparene","compound.1","alpha.chamigrene",
               "widdrol","cedrol","beta.acorenol","alpha.acorenol","gamma.eudesmol",
               "beta.eudesmol","alpha.eudesmol","cedr.8.en.13.ol","cedr.8.en.15.ol",
               "compound.2","thujopsenal")


juniper_tidy <- juniper_dat %>% 
  pivot_longer(cols = all_of(chemicals),
               names_to = "ChemicalID",
               values_to = "Concentration")
```

The structure of the newly created data set is now longer and much more tidy.
```{r}
str(juniper_tidy)
```

## 4. Make the given graph.
Now we make a plot faceted by ChemicalID that shows the Concentration of each chemical over time (YearsSinceBurn).
```{r}
ggplot(data = juniper_tidy, aes(x = YearsSinceBurn, y = Concentration)) +
  geom_smooth() +
  facet_wrap(~ ChemicalID, scales = "free") +
  theme_minimal() +
  labs(x = "YearsSinceBurn", y = "Concentration")

```

## 5. Use a generalized linear model to find which chemicals show concentrations that are significantly (significant, as in P < 0.05) affected by “Years Since Burn”.

```{r}

glm_signif_results <- juniper_tidy %>% 
  group_by(ChemicalID) %>% 
  nest() %>% 
  mutate(model = map(data, function(new_df) glm(Concentration ~ YearsSinceBurn, data = new_df))) %>% 
  mutate(tidy_model = map(model, tidy)) %>% 
  unnest(tidy_model) %>% 
  filter(p.value < 0.05) %>% 
  select(ChemicalID, term, estimate, std.error, statistic, p.value)
  
print(glm_signif_results, n = Inf)
```




