---
title: "HATCHER_Skills_Test_3"
output: html_document
date: "2025-04-21"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1. Load and clean FacultySalaries_1995.csv file and Re-create the given graph:

```{r include=FALSE}
# load in the needed libraries
library(tidyverse)
library(janitor)
library(ggplot2)

# load in the data
fac_dat <- read.csv("FacultySalaries_1995.csv")

# take a look at the data structure
# str(fac_dat)
```

```{r}
# clean the data by reshaping it to the long format
faculty_long <- fac_dat %>% 
  pivot_longer(
    cols = c(AvgFullProfSalary, AvgAssocProfSalary, AvgAssistProfSalary),
    names_to = "Rank",
    values_to = "Salary") %>% 
  mutate(Rank = case_when(
    Rank == "AvgFullProfSalary" ~ "Full",
    Rank == "AvgAssocProfSalary" ~ "Assoc",
    Rank == "AvgAssistProfSalary" ~ "Assist",
    TRUE ~ Rank))

# make rank, tier, and state variables into factors for plotting
faculty_long$Rank <- as.factor(faculty_long$Rank)
faculty_long$Tier <- as.factor(faculty_long$Tier)
faculty_long$State <- as.factor(faculty_long$State)
```

```{r}
# graph the plot
faculty_long_plot <- filter(faculty_long, Tier != "VIIB")

ggplot(faculty_long_plot, aes(x = Rank, y = Salary, fill = Rank)) +
  geom_boxplot() +
  facet_wrap(~ Tier) +
  labs(y = "Salary", x = "Rank") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 60))
```

# 2. Build an ANOVA model and display the summary output in your report.

```{r}
# build the model
anova_mod <- aov(data = faculty_long, Salary ~ State + Tier + Rank)

summary(anova_mod)
```

##### Based on the ANOVA model above, we can conclude that all three variables (State, Tier, and Rank) have effects on faculty salaries that are statistically significant.  
  
##### Since Rank has the largest Sum Sq and F value, it has the largest effect on the variation in salaries.

# 3. Load and look at "Juniper_Oils.csv" data and then tidy it.

```{r}
juniper_dat <- read.csv("Juniper_Oils.csv")

# str(juniper_dat)
```

```{r}
# make a vector of the chemical compounds
chemicals <- c("alpha.pinene","para.cymene","alpha.terpineol","cedr.9.ene",
               "alpha.cedrene","beta.cedrene","cis.thujopsene","alpha.himachalene",
               "beta.chamigrene","cuparene","compound.1","alpha.chamigrene",
               "widdrol","cedrol","beta.acorenol","alpha.acorenol","gamma.eudesmol",
               "beta.eudesmol","alpha.eudesmol","cedr.8.en.13.ol","cedr.8.en.15.ol",
               "compound.2","thujopsenal")

```

```{r}
# tidy up the chemical data while keeping the rest

juniper_tidy <- juniper_dat %>% 
  pivot_longer(cols = all_of(chemicals),
               names_to = "ChemicalID",
               values_to = "Concentration")

```

# 4. Make the given graph.

```{r}
# make a plot that shows the Concentration of each chemical over time since the YearsSinceBurn faceted by ChemicalID
ggplot(data = juniper_tidy, aes(x = YearsSinceBurn, y = Concentration)) +
  geom_smooth() +
  facet_wrap(~ ChemicalID, scales = "free") +
  theme_minimal() +
  labs(x = "YearsSinceBurn", y = "Concentration")

```

# 5. Use a generalized linear model to find which chemicals show concentrations that are significantly (significant, as in P < 0.05) affected by “Years Since Burn”.

```{r}
library(broom)
glm_signif_results <- juniper_tidy %>% 
  group_by(ChemicalID) %>% 
  nest() %>% 
  mutate(model = map(data, function(new_df) glm(Concentration ~ YearsSinceBurn, data = new_df))) %>% 
  mutate(tidy_model = map(model, tidy)) %>% 
  unnest(tidy_model) %>% 
  filter(p.value < 0.05) %>% 
  select(ChemicalID, term, estimate, std.error, statistic, p.value)
  
glm_signif_results
```




